The Astropy Project: Building an inclusive, open science project and status of the v2.0 software

### 2018-01-08 main.tex
Added some clarifications on io.ascii compression support and fast reader capabilities

\section{Introduction} \label{sec:intro}
All modern astronomical research makes use of software in some way.
Astronomy as a field has thus long supported the development of software tools
for astronomical tasks, such as scripts that enable individual scientific
research, software packages for small collaborations, and data reduction
pipelines for survey operations.
Some software packages are or were supported by large institutions and are
intended for a wide range of users.
These packages therefore typically provide some level of documentation and user
support or training.
Other packages are developed by individual researchers or research groups and
are then typically used by smaller groups for more domain-specific purposes.
For both packages meant for wider distribution and for scripts specific to
particular research projects, a library that addresses common astronomical tasks
simplifies the software development process.
The users of such a library then also benefit from a community and ecosystem
built around a shared foundation.
The \astropy project has grown to become this community for \python astronomy
software, and the \astropypkg core package a feature-rich \python library.

@@

The \package{astropy.io} subpackage provides support for reading and writing
data to a variety of ASCII and binary file formats, such as a wide range of
ASCII data table formats, FITS, and VOTable.
It also provides a unified interface for reading and writing data with these
different formats using the \package{astropy.table} subpackage.
For many common cases, this simplifies the process of file input and output (I/O) and
reduces the need to master the separate details of all the I/O packages within
\astropypkg.

@@

The \package{astropy.io.ascii} subpackage now includes a significantly faster
Cython/C engine for reading and writing ASCII files. This is available for most
of the common formats.  On average, the new engine is about 4 to 5 times faster
than the corresponding pure-\python implementation and is often comparable to
the speed of the \package{pandas} \citep{pandas} ASCII file interface.  The
fast reader has a parallel processing option that allows harnessing multiple
cores for input parsing to achieve even greater speed gains.  By default,
\texttt{read()} and \texttt{write()} will attempt to use the fast Cython/C engine
when dealing with compatible formats. Certain features of the full read
/ write interface are unavailable in the fast version, in which case the
pure-\python version will automatically be used.

@@

The following individuals would like to recognize support for their personal contributions.
HMG was supported by the National Aeronautics and Space Administration through the 
Smithsonian Astrophysical Observatory contract SV3-73016 to MIT for Support of the Chandra X-Ray Center, which is
operated by the Smithsonian Astrophysical Observatory for and on behalf of the National Aeronautics Space 
Administration under contract NAS8-03060. JTV was supported by the UW eScience Institute, via grants from the Moore
Foundation, the Sloan Foundation, and the Washington Research Foundation. SMC acknowledges the National Research
Foundation of South Africa. TLA was supported by NASA contract NAS8-03060. Support for E.J.T. was provided by NASA
through Hubble Fellowship grant No. 51316.01 awarded by the Space Telescope Science Institute, which is operated by
the Association of Universities for Research in Astronomy, Inc., for NASA, under contract NAS 5-26555, as well as a 
Giacconi Fellowship. MB was supported by the FONDECYT regular project 1170618 and the MINEDUC-UA projects codes 
ANT 1655 and ANT 1656. DH was supported through the SFB 881 "The Milkyway System" by the German Research Foundation 
(DFG). WEK was supported by an ESO Fellowship. C.M. is supported by NSF grant AST-1313484. SP was supported by 
grant AYA2016-75808-R (FEDER) issued by the Spanish government. JEHT was supported by the Gemini Observatory, which
is operated by the Association of Universities for Research in Astronomy, Inc., on behalf of the international 
Gemini partnership of Argentina, Brazil, Canada, Chile, and the United States of America.

>>>>

\section{Introduction} \label{sec:intro}
All astronomical research makes use of software in some way.
Astronomy as a field has thus long supported the development of software tools
for astronomical tasks, such as scripts that enable individual scientific
research, software packages for small collaborations, and data reduction
pipelines for survey operations.
Some software packages are or were supported by large institutions and are
intended for a wide range of users.
These packages therefore typically provide some level of documentation and user
support or training.
Other packages are developed by individual researchers or research groups and
are then typically used by smaller groups for more domain-specific purposes.
For both packages meant for wider distribution and for scripts specific to
particular research projects, a library that addresses common astronomical tasks
simplifies the software development process.
The users of such a library then also benefit from a community and ecosystem
built around a shared foundation.
The \astropy project has grown to become this community for \python astronomy
software, and the \astropypkg core package a feature-rich \python library.

@@

The \package{astropy.io} subpackage provides support for reading and writing
data to a variety of ASCII and binary file formats, such as a wide range of
ASCII data table formats, FITS, and VOTable.
It also provides a unified interface for reading and writing data with these
different formats using the \package{astropy.table} subpackage.
For many common cases, this simplifies the process of file input and output (I/O) and
reduces the need to master the separate details of all the I/O packages within
\astropypkg. The file interface allows transparent compression of
the \texttt{gzip}, \texttt{bzip2} and \texttt{lzma} (\texttt{.xz})
formats; for the latter two if the Python installation was compiled
with support the respective libraries.

@@

The \package{astropy.io.ascii} subpackage now includes a significantly faster
Cython/C engine for reading and writing ASCII files. This is available for most
of the common formats. It also offers some additional features like parsing
of different exponential notation styles, such as commonly produced
by Fortran programs.  On average, the new engine is about 4 to 5 times faster
than the corresponding pure-\python implementation and is often comparable to
the speed of the \package{pandas} \citep{pandas} ASCII file interface.  The
fast reader has parallel processing option that allows harnessing multiple
cores for input parsing to achieve even greater speed gains.  By default,
\texttt{read()} and \texttt{write()} will attempt to use the fast Cython/C engine
when dealing with compatible formats. Certain features of the full read
/ write interface are unavailable in the fast version, in which case the
reader will by default fall back automatically to the pure-\python version.

@@

The following individuals would like to recognize support for their personal contributions.
HMG was supported by the National Aeronautics and Space Administration through the 
Smithsonian Astrophysical Observatory contract SV3-73016 to MIT for Support of the Chandra X-Ray Center, which is
operated by the Smithsonian Astrophysical Observatory for and on behalf of the National Aeronautics Space 
Administration under contract NAS8-03060. JTV was supported by the UW eScience Institute, via grants from the Moore
Foundation, the Sloan Foundation, and the Washington Research Foundation. SMC acknowledges the National Research
Foundation of South Africa. TLA was supported by NASA contract NAS8-03060. Support for E.J.T. was provided by NASA
through Hubble Fellowship grant No. 51316.01 awarded by the Space Telescope Science Institute, which is operated by
the Association of Universities for Research in Astronomy, Inc., for NASA, under contract NAS 5-26555, as well as a 
Giacconi Fellowship. MB was supported by the FONDECYT regular project 1170618 and the MINEDUC-UA projects codes 
ANT 1655 and ANT 1656. DH was supported through the SFB 881 ``The
Milky Way System'' by the German Research Foundation 
(DFG). WEK was supported by an ESO Fellowship. C.M. is supported by NSF grant AST-1313484. SP was supported by 
grant AYA2016-75808-R (FEDER) issued by the Spanish government. JEHT was supported by the Gemini Observatory, which
is operated by the Association of Universities for Research in Astronomy, Inc., on behalf of the international 
Gemini partnership of Argentina, Brazil, Canada, Chile, and the United States of America.
